State and Parameter Estimation for Spatio-Temporal Neural Fields

### 2010-05-09 ltx/Journal of Neuroengineering/StateAndParameterEstimationNeuralFields_JoNE.tex

@@

   	\caption{Example of experimental design and model output. \textbf{a}. Example of the neural field with sensors. The centres of the sensors are shown by crossed and the sensor widths ($1/2$ height of Gaussian) are shown by the circles. \textbf{b} Example of data generated by the model through the observations. \textbf{c}. Power spectral density of the data generated by an observation showing the typical $1/f$ characteristic of iEEG.} 

@@

runs.}

@@

\begin{figure*}
	\begin{center}
\subfigure[][]{\includegraphics[width=0.3\textwidth]{./Graph/BasisChar.pdf}}
\subfigure[][]{\includegraphics[width=0.3\textwidth]{./Graph/SensorChar.pdf}}
\subfigure[][]{\includegraphics[width=0.3\textwidth]{./Graph/ObservationFrequencyResponse.pdf}}
	\end{center}
	\caption{a:Field basis function frequency response; b:Sensor frequency response ; c: The average FFT of the observed field.} 
\label{fig:FreqAnalysis} 
\end{figure*}
Following this, the distance between basis function, $\Delta_b$, was set to $2.5$ and the basis function width, $\sigma_{\phi}^2$, was set to 3.6, 3.16 mm full width at half maximum. Analysis said 3.9 from Parham's derivation, but 3.6 still overlaps and allows more high frequency details to be represented.

\subsection{State and Parameter Estimates} To demonstrate the performance of the parameter and state estimation, $250$ realisations of the data were generated. Each realisation consisted of 200 ms of data and the estimation was applied to the final 200 ms, allowing the model to stabilise. The initial state and parameters were unknown to the estimator. The distribution of the parameters and the resulting kernel are shown in figure~\ref{fig:Parameters} and figure~\ref{fig:KernelEstimates} respectively. Although a wide distribution can be seen over the parameter estimates the ratio between them remans in a narrow range as shown in figure~\ref{fig:ParametersRatio}. The state estimates for one realization were evaluated by comparing the field reconstruction to the true field using the mean over time of the Root Mean Square Error (RMSE) and standard deviation of the RMSE giving $0.19\pm 0.01$. The RMSE of the field estimate over time is also shown in figure~\ref{fig:RMSE}. One snapshot for the original and the field estimate and their corresponding error are illustrated in figure~\ref{fig:FieldEstimate}.
\begin{figure}
\includegraphics[width=0.5\textwidth]{./Graph/KernelEstimate_KernelFreqResponse.pdf}
\caption{Mean kernel estimates over 250 runs ;  true and estimated kernels are shown by solid
and dashed line respectively, the gray area shows 95\% confedence interval; top: spatial domain; bottom: frequency domain}
\label{fig:KernelEstimates}
\end{figure}
\begin{figure}[th]
\subfigure[][]{\includegraphics[width=0.24\textwidth]{./Graph/theta0_theta1_ratio.pdf}}
\subfigure[][]{\includegraphics[width=0.24\textwidth]{./Graph/theta0_theta2_ratio.pdf}}\\
\caption{Distribution of the parameters ratios over 250 runs.}
\end{figure}
\label{fig:ParametersRatio}
\begin{figure}[th]
\subfigure[][]{\includegraphics[width=0.24\textwidth]{./Graph/FFTTrueField.pdf}}\label{fig:FieldFFT}
\subfigure[][]{\includegraphics[width=0.24\textwidth]{./Graph/FFTEstimatedField.pdf}}\label{fig:EstimatedFieldFFT}\\
\caption{The average FFT of the hidden field; a: True field; b: Estimated field.}
\label{fig:FFTTrueEstimate}
\end{figure}

>>>>

	$\Sigma$ & & field disturbance covariance \\
	sensor parameters & & \\
	$\Delta$ & 0.5 & space step \\
	$T_s$ & 1 & ms time step \\ 

@@

   	\caption{Example of experimental design and model output. \textbf{a}. Example of the neural field with sensors. The centres of the sensors are shown by crosses. \textbf{b} Example of data generated by the model through the observations. \textbf{c}. Power spectral density of the data generated by an observation showing the typical $1/f$ characteristic of iEEG.} 

@@

runs. True parameters are $\zeta=100$, $\theta_0 = 10$, $\theta_1 = -8$, $\theta_3 = 0.5$. Grey bars show distributions with wider basis functions and blue bars show improved parameter distributions with narrower basis functions.}

@@

\begin{figure*}
	\begin{center}
\subfigure[][]{\includegraphics[width=0.3\textwidth]{./Graph/BasisChar.pdf}}
\subfigure[][]{\includegraphics[width=0.3\textwidth]{./Graph/SensorChar.pdf}}
\subfigure[][]{\includegraphics[width=0.3\textwidth]{./Graph/ObservationFrequencyResponse.pdf}}
	\end{center}
	\caption{a:Field basis function frequency response (projected to a 1D plot because it is symmetric); b:Sensor frequency response (projected to a 1D plot because it is symmetric); c: The average FFT of the observed field.} 
\label{fig:FreqAnalysis} 
\end{figure*}
Following this, the distance between basis function, $\Delta_b$, was set to $2.5$ and the basis function width, $\sigma_{\phi}^2$, was set to 3.6, 3.16 mm full width at half maximum. Analysis said 3.9 from Parham's derivation, but 3.6 still overlaps and allows more high frequency details to be represented.

\subsection{State and Parameter Estimates} To demonstrate the performance of the parameter and state estimation, $250$ realisations of the data were generated. Each realisation consisted of 200 ms of data and the estimation was applied to the final 200 ms, allowing the model to stabilise. The initial state and parameters were unknown to the estimator. The distribution of the parameters and the resulting kernel are shown in figure~\ref{fig:Parameters} and figure~\ref{fig:KernelEstimates} respectively. Although a wide distribution can be seen over the parameter estimates the ratio between them remans in a narrow range as shown in figure~\ref{fig:ParametersRatio}. The state estimates for one realization were evaluated by comparing the field reconstruction to the true field using the mean over time of the Root Mean Square Error (RMSE) and standard deviation of the RMSE giving $0.19\pm 0.01$. The RMSE of the field estimate over time is also shown in figure~\ref{fig:RMSE}. One snapshot for the original and the field estimate and their corresponding error are illustrated in figure~\ref{fig:FieldEstimate}.
\begin{figure}
\includegraphics[width=0.5\textwidth]{./Graph/KernelEstimate_KernelFreqResponse.pdf}
\caption{Mean kernel estimates over 250 runs ;  true and estimated kernels are shown by solid
and dashed line respectively, the grey area shows 95\% confidence interval; Top: spatial domain; Bottom: frequency domain}
\label{fig:KernelEstimates}
\end{figure}
\begin{figure}[th]
\subfigure[][]{\includegraphics[width=0.24\textwidth]{./Graph/theta0_theta1_ratio.pdf}}
\subfigure[][]{\includegraphics[width=0.24\textwidth]{./Graph/theta0_theta2_ratio.pdf}}\\
\caption{Distribution of the parameters ratios over 250 runs.}
\end{figure}
\label{fig:ParametersRatio}
\begin{figure}[th]
\subfigure[][]{\includegraphics[width=0.24\textwidth]{./Graph/FFTTrueField.pdf}}\label{fig:FieldFFT}
\subfigure[][]{\includegraphics[width=0.24\textwidth]{./Graph/FFTEstimatedField.pdf}}\label{fig:EstimatedFieldFFT}\\
\caption{The average FFT of the neural field; a: True field; b: Estimated field (showing the limited bandwidth).}
\label{fig:FFTTrueEstimate}
\end{figure}
